---
title: "Assignment2-Wang-Oya"
subtitle: "Due at 11:59pm on October 3."
author: "Akari Oya, Zhouer Wang"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#load libraries
library(tidyverse)
library(gtrendsR)
library(censusapi)
```

## Github link = https://github.com/ZuorW/SURV727.git

\
In this assignment, you will pull from APIs to get data from various data sources and use your data wrangling skills to use them all together. You should turn in a report in PDF or HTML format that addresses all of the questions in this assignment, and describes the data that you pulled and analyzed. You do not need to include full introduction and conclusion sections like a full report, but you should make sure to answer the questions in paragraph form, and include all relevant tables and graphics.

Whenever possible, use piping and `dplyr`. Avoid hard-coding any numbers within the report as much as possible.\

## Pulling from APIs

Our first data source is the Google Trends API. Suppose we are interested in the search trends for `crime` and `loans` in Illinois in the year 2020. We could find this using the following code:

```{r}
res <- gtrends(c("crime", "loans"), 
               geo = "US-IL", 
               time = "2020-01-01 2020-12-31", 
               low_search_volume = TRUE)
plot(res)
```

Answer the following questions for the keywords "crime" and "loans".

-   Find the mean, median and variance of the search hits for the keywords.

```{r}
# Check what the data looks like
res$interest_over_time %>% 
  head()

# Take interest over time data and group data into crime and loans in separate rows
# Also compute the mean, sd, variance of each keyword
res$interest_over_time %>% 
  group_by(keyword) %>% 
  summarize(mean_hits = mean(hits),
            median = median(hits),
            var_hits = var(hits))
```

The keyword `crime` had a mean search hit of 55.230 with a median of 53.5 and a variance of 78.965 The keyword `loans` had a mean search hit of 66.942 with a median of 66.5 and a variance of 99.624\

-   Which cities (locations) have the highest search frequency for `loans`? Note that there might be multiple rows for each city if there were hits for both "crime" and "loans" in that city. It might be easier to answer this question if we had the search hits info for both search terms in two separate variables. That is, each row would represent a unique city.

```{r}
# Check what the data looks like
res$interest_by_city %>% 
  head()

# Reshape data from long to wide format using keywords
wide <- 
  res$interest_by_city %>% 
  pivot_wider(names_from = keyword,
              values_from = hits)

# Check for duplicates in location to see if reshaping worked and show in table if any exist
wide %>%
  add_count(location) %>%
  filter(n>1) %>%
  distinct()

# Sort loans column in descending order
wide %>%   
  arrange(desc(loans)) %>% 
  head()
```
The cities Granville, Alorton, and Bement have the highest search frequency for `loans`.\

-   Is there a relationship between the search intensities between the two keywords we used?

```{r}
# Run Pearson correlation test
cor.test(wide$crime, wide$loans)
```

While `loans` had a higher mean search frequency over time, there does not seem to be a large difference between the search frequency of `crime`. However, patterns can be seen in the plot. The two keywords seems to have an inverse relationship where search frequencies for `loans` are high when `crime` is low in the first peak/dip around April 2020. However, the pattern fades after around July 2020. We tested the relationship between the variables for interest by city to properly with a Pearson correlation test. The test suggests that there is no significant correlation (r = -0.115, p = .72).\

Repeat the above for keywords related to covid. Make sure you use multiple keywords like we did above. Try several different combinations and think carefully about words that might make sense within this context.

```{r}
# Create another dataset with new keywords
res_2 <- gtrends(c("masks", "deaths"), 
               geo = "US-IL", 
               time = "2020-01-01 2020-12-31", 
               low_search_volume = TRUE)
plot(res_2)
```

```{r}
res_2 %>% 
  head()

# Compute the mean, standard deviation, and variance of search hits per keyword
res_2$interest_over_time %>% 
  group_by(keyword) %>% 
  summarize(mean_hits = mean(hits),
            median_hits = median(hits),
            var_hits = var(hits))
```

```{r}
# Check data
res_2$interest_by_city %>% 
  arrange(desc(location)) %>% 
  glimpse()

# There is a duplicate location name (Willowbrook) and pivot_wider cannot be run
# Check for duplicate case using code given in error message
res_2$interest_by_city %>% 
  group_by(location, geo, gprop, keyword) %>% 
  summarise(n = n(), .groups = "drop") %>% 
  filter(n > 1L) #shows duplicate location names

# Remove it from the data for the sake of this assignment
res2_nodup <- 
  res_2$interest_by_city %>% 
  filter(location != "Willowbrook")

# Reshape data from long to wide format using keywords
wide_2 <- 
  res2_nodup %>% 
  pivot_wider(names_from = keyword,
           values_from = hits)

# Sort masks column in descending order
wide_2 %>%   
  arrange(desc(masks)) %>% 
  head()

# Run Pearson correlation test
cor.test(wide_2$masks, wide_2$deaths)
```

We used the keywords `masks` and `deaths`. The search frequency for `masks` over time had a mean of 32.712 with a standard deviation of 23.99 and a variance of 575.543 The search frequency for `deaths` over time had a mean of 31.365 with a standard deviation of 22.03 and a variance of 485.334 Both keywords have a similar mean and have high variances. In the plot, the frequencies have a similar shape in the first half of 2020, but the pattern becomes unclear in the second half. The initial spike in search hits of both keywords understandably corresponds to near the beginning of the pandemic when everyone was required to wear masks. We conducted a Pearson correlation test to see if the search frequencies of the two keywords have a relationship. The test revealed that there is no significant correlation (r = 0.09, p = .5946).
	

## Google Trends + ACS

Now lets add another data set. The `censusapi` package provides a nice R interface for communicating with this API. However, before running queries we need an access key. This (easy) process can be completed here:

<https://api.census.gov/data/key_signup.html>

Once you have an access key, store this key in the `cs_key` object. We will use this object in all following API queries.

```{r}
#| eval: false
cs_key <- "7b1cc9af0a42634e3ba57f9a8f5d0098cdedc5e4"
```

In the following, we request basic socio-demographic information (population, median age, median household income, income per capita) for cities and villages in the state of Illinois.

```{r}
#| eval: false

acs_il <- getCensus(name = "acs/acs5",
                    vintage = 2020, 
                    vars = c("NAME", 
                             "B01001_001E", 
                             "B06002_001E", 
                             "B19013_001E", 
                             "B19301_001E"), 
                    region = "place:*", 
                    regionin = "state:17",
                    key = cs_key)
head(acs_il)
```

Convert values that represent missings to NAs.

```{r}
#| eval: false

acs_il[acs_il == -666666666] <- NA
```

Now, it might be useful to rename the socio-demographic variables (`B01001_001E` etc.) in our data set and assign more meaningful names.

```{r}
#| eval: false
acs_il <-
  acs_il %>%
  rename(pop = B01001_001E, 
         age = B06002_001E, 
         hh_income = B19013_001E, 
         income = B19301_001E)
```

It seems like we could try to use this location information listed above to merge this data set with the Google Trends data. However, we first have to clean `NAME` so that it has the same structure as `location` in the search interest by city data. Add a new variable `location` to the ACS data that only includes city names.

```{r}
# Check headers
acs_il %>% 
  head()

# Create new location variable
no_village <- gsub(' village, Illinois', '', acs_il$NAME) #remove "village, IL" from NAME and store
no_cityvill <- gsub(' city, Illinois', '', no_village) #take above and remove remaining "city, IL"
acs_with_loc <- 
  acs_il %>% 
  mutate(location = no_cityvill) #add new variable with only city names

acs_with_loc %>% 
  head()
```

## Answer the following questions with the "crime" and "loans" Google trends data and the ACS data.

-   First, check how many cities don't appear in both data sets, i.e. cannot be matched. Then, create a new data set by joining the Google Trends and the ACS data. Keep only cities that appear in both data sets.

```{r}
# Check how many cities cannot be matched from ACS data
acs_with_loc %>% 
  anti_join(wide, by = "location") %>% 
  count() #show number of rows (cities)

# Merge ACS to gtrends data by city only keeping cases that match
merged <-
  wide %>% 
<<<<<<< HEAD
  inner_join(acs_with_loc, by = "location")
=======
  left_join(acs_with_loc, by = "location")
# should we use left_join or inner_join here?
>>>>>>> 8b14667251440d22588929dc29aab0dc0e141c4c

merged %>% 
  head()

 uni_acs_il <- acs_with_loc %>%
       distinct(location)
```

\
- Compute the mean of the search popularity for both keywords for cities that have an above average median household income and for those that have an below average median household income. When building your pipe, start with creating the grouping variable and then proceed with the remaining tasks. What conclusions might you draw from this?
```{r}
# If household income is greater than its median, name group as above average, if not, name group as above average
# Then compute mean by group
merged %>% 
  group_by(
    hhinc_med = 
      ifelse(hh_income > mean(hh_income, na.rm = TRUE), 
                       "above", "below")) %>% 
                       summarize(mean_crime = mean(crime, na.rm = TRUE),
                       mean_loans = mean(loans, na.rm = TRUE)) #code doesn't work if I don't use na.rm = T
```
For cities that have an above average median household income, the search popularity of `crime` was 42.778 and 33.115 for `loans`. For cities that have a below average median household income, the search popularity of `crime` was 45.032 and 39.429 for `loans`. Those in cities with below average household income had a higher search rate for both keywords. We conclude that crime rates may be higher in below average cities which may lead to more search hits, and that people in these cities may search for `loans` more because more people in these cities may take out loans to support their lives due to a lower financial status. \

-   Is there a relationship between the median household income and the search popularity of the Google trends terms? Describe the relationship and use a scatterplot with `qplot()`.
```{r}
# Plot for crime
qplot(hh_income, crime, data = merged)

# Correlation test
cor.test(merged$hh_income, merged$crime)

# Plot for loans
qplot(hh_income, loans, data = merged)

# Correlation test
cor.test(merged$hh_income, merged$loans)
```

In the plot for the median household income and the search popularity of `crime`, much of the data is gathered in the lower half of the median household income but there is no clear pattern. A Pearson correlation test supports this by showing that there is no correlation between the two variables (r = -0.108, p = .352). On the other hand, the plot for `loans` shows a clear pattern in which higher search hits are centered around the lower end of median household income, suggesting a relationship between the two variables. We tested this relationship using a Pearson correlation test. There was a significant correlation (r = -0.344, p < .001).

## Repeat the above steps using the covid data and the ACS data.
```{r}
# Check how many cities cannot be matched from ACS data to covid gtrends data
acs_with_loc %>% 
  anti_join(wide_2, by = "location") %>% 
  count() #show number of rows (cities)

# Merge ACS to gtrends data by city only keeping cases that match
merged_2 <-
  wide_2 %>% 
  left_join(acs_with_loc, by = "location")

merged_2 %>% 
  head()
```

```{r}
# If household income is greater than its median, name group as above average, if not, name group as above average
# Then compute mean by group
merged_2 %>% 
  group_by(
    hhinc_med = 
      ifelse(hh_income > median(hh_income, na.rm = TRUE), 
                       "above", "below")) %>% 
                       summarize(mean_masks = mean(masks, na.rm = TRUE),
                       mean_deaths = mean(deaths, na.rm = TRUE)) #code doesn't work if I don't use na.rm = True
```

```{r}
# Plot for masks
qplot(hh_income, masks, data = merged_2)

# Correlation test
cor.test(merged_2$hh_income, merged_2$masks)

# Plot for deaths
qplot(hh_income, deaths, data = merged_2)

# Correlation test
cor.test(merged_2$hh_income, merged_2$deaths)
```
The Pearson correlation test shows that `masks` have a relationship with median household income (r = 0.324, p < .001). The data for `mask` search hits in the plot has less outliers with most of the data points gathering around the lower side of income. On the other hand, `deaths` did not have a relationship with median household income (r = 0.124, p = .209). This coincides with the data points in the plot for `deaths` being more spread out. Notably, people with lower household income who may be more at risk of being infected or spreading COVID-19 due to their socioeconomic status, may have searched for `masks` frequently to buy or make them. 
